---
title: "Reproducible Research Project"
author: "Clara"
date: "November 17, 2018"
output: html_document
---

##**Loading and preprocessing the data**
``` {r echo = TRUE}
activity <- read.csv("./activity.csv", header = TRUE)
activity[,2] <- as.Date(activity$date)
head(activity)
```
##**What is the mean total number of steps taken per day?**

1. Total number of steps taken per day
```{r echo = TRUE}
steps_1 <- with(activity, tapply(steps, date, sum, na.rm = TRUE))
```

2. Histogram of total number of steps per day
```{r echo=TRUE}
hist(steps_1, col = "blue", xlab = "Sum of Steps", ylab = "Frequency", main = " Total Number of Steps Per Day")
```


3. Mean and Median of total number of steps taken per day
```{r echo = TRUE, results='asis'}
print(mean_steps <- mean(steps_1))
print(median_steps <- median(steps_1))
```

The average steps taken per day is `r mean_steps` and the median of total number of steps taken is `r median_steps`

##**What is the average daily activity pattern? **

1. Make a time-series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo = TRUE}
average_steps <- with(activity, tapply(steps, interval, mean, na.rm = TRUE))
intervals <- unique(activity$interval)
new <- data.frame(cbind(average_steps,intervals))
plot(new$intervals, new$average_steps, type = "l", xlab = "Intervals", ylab = "Average Steps", main = "Average Steps Per Interval")
```

2. Which 5-minute interval contains the maximum number of steps?

```{r echo = TRUE, results='asis'}
index <- which.max(new$average_steps)
max <- new[index,2]
```
The 5-minute interval that contains the maximum the maximum number of steps is `r max`

##**Inputing missing values**

1. Total number of missing values in the dataset.

```{r echo = TRUE, results='asis'}
missing_values <- sum(is.na(activity$steps))
```
The sum of missing values in the dataset is `r missing_values`

2. Develop a strategy for filling in all missing values.

**Strategy:** Calculate the average of average steps per day across all dates in the dataset, ignoring NA values. Then use the resulting values in place of NAs.

3. Create a new dataset that is equal to the original dataset but with the NA values filled in.

```{r echo = TRUE}
index <- which(is.na(activity$steps))
l <- length(index)
steps_avg <- with(activity,tapply(steps, date, mean, na.rm = TRUE))
na <- mean(steps_avg, na.rm = TRUE)
for (i in 1:l) {
    activity[index[i],1] <- na
} 
```
Are NAs properly filled?

```{r echo = TRUE}
check_NAs <- sum(is.na(activity$steps))
```
Since check_NAs is equal to `r check_NAs`, all NAs are properly filled.

Look at new data set
``` {r}
str(activity)
head(activity)
```

4. Histogram of steps taken each day.
``` {r echo = TRUE}
steps_2 <- with(activity, tapply(steps, date, sum, na.rm = TRUE))
hist(steps_2, col = "purple", xlab = "Total Steps",ylab = "Frequency", main = "Total Steps Per Day")
```

Calculate mean and median of total steps taken per day

```{r echo = TRUE, results='asis'}
print(mean_steps_2 <- mean(steps_2))
print(median_steps_2 <- median(steps_2))
```

The average steps taken per day is `r mean_steps_2` and the median of total number of steps taken is `r median_steps_2`

**Impact of inputing missing values on the estimates of total daily number of steps:** The mean and median of the toal steps are equal.

##**Are there differences in activity patterns between weekends and weekdays?**

1. Create the weekday and weekend variables.
``` {r echo = TRUE}
library(dplyr)

activity_mod <- mutate(activity, day = ifelse(weekdays(activity$date) == "Saturday" | weekdays(activity$date) == "Sunday", "weekend", "weekday"))

activity_mod$day <- as.factor(activity_mod$day)

str(activity_mod)
```

2. Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

``` {r}
act_wknd<-subset(activity_mod,as.character(activity_mod$day)=="weekend")
act_wkdy<-subset(activity_mod,as.character(activity_mod$day)=="weekday")

steps_wknd<-with(act_wknd,tapply(steps,interval,mean,na.rm=TRUE))
steps_wkdy<-with(act_wkdy,tapply(steps,interval,mean,na.rm=TRUE))

int_wknd<-unique(act_wknd$interval)
int_wkdy<-unique(act_wkdy$interval)

new_wknd<-data.frame(cbind(steps_wknd,int_wknd))
new_wkdy<-data.frame(cbind(steps_wkdy,int_wkdy))

par(mfrow=c(2,1),mar=c(4,4,2,1))

plot(new_wknd$int_wknd,new_wknd$steps_wknd,type = "l",xlab = "Intervals",
     ylab = "Average Steps",main = "Weekend")

plot(new_wkdy$int_wkdy,new_wkdy$steps_wkdy,type = "l",xlab = "Intervals",
     ylab = "Average Steps",main = "Weekday")
```
